Ancestry linkage in hybrid zones

### 2016-02-14 drafts/hybrid_zone.tex
got to the end of methods

\section{Introduction}
The process of speciation commonly involves allopatric divergence in relative isolation \citep{Coyne2004}.  
When formerly isolated populations come back into contact, \yb{before reproductive isolation is complete, they can exchange genes.}  
\yb{Thus, ancestry originating from the two diverged populations spreads from their point of contact across a geographic hybrid zone --- establishing a gradient of ancestry across geography, \citep{Barton1985}}. 
%In addition to steep clines in ancestry,
\yb{The recent exchange of genetic material between diverged populations also leaves a distinct pattern of variation across the genome, in which long tracts of ancestry from each  are broken up by historical recombination events.}  
\yb{To the extent that the genetic variation can be exchanged between species via hybridization, both of the geographic clines in ancestry, and extreme linkage disequilibrium in ancestry (ancestry-LD) diminish over time.} 
However, subsets of the genome that are under selection can maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}. 
The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986}. 
\yb{However the expected length of ancestral tracts around selected loci in admixing populations is poorly understood.}

Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place every generation \yb{(CITE Fisher; Chapman \& Thomson; Barid et al)}.   
The expected distribution of lengths of unbroken ancestry have been previously described under neutral conditions \cite[e.g.][]{Gravel2012,Sedghifar2015}.  
Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection. 
Selection against migrant \yb{alleles} in a hybrid zone results in reduced frequencies of foreign ancestry surrounding targets of selection \yb{(CITE CLINES)}. 
The effect on patterns of ancestral block lengths is, however, less well understood. 
Because of rapid removal from the population, we expect to observe \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone. 
\alisa{While empirical patterns of block length in hybrid zones have not been well studied, the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones. } %GREAT SENTENCE!!

The extensive history of population genetic modeling of hybrid zones has focused largely on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}.  
These models predict cline width to be narrowest \yb{(i.e. the transition in allele frequency is the steepest)} at loci close to selected loci, and cline width at less tightly linked loci to be wider. %and more heavily influenced by demography.} %hybrid zone age and rates of gene flow. 
\yb{The focus on cline width provide an expectation for how selection shapes the covariation in ancestry at a single locus over \emph{geographic space}, our aim is to provide an expectation for how selection shapes the covariation in ancestry at loci over \emph{genomic space.}}  
%As denser genotype data becomes more readily available, genome-wide patterns of linkage can be utilized in the analysis of gene hybrid zones.

@@

\paragraph{Lineages at the selected locus}
Now suppose that the frequency profile of the selected allele $p(x,t)$ is given.
We will understand how lineages of sampled individuals behave,
both at and away from the selected site,
by following lineages as they move between backgrounds,
as done for instance in \citet{selectioncoal,ralph2015patchy}.
Consider the collection of $A$ alleles found at location $x$.
The expected number of such alleles produced by location $y$ is proportional to 
the number of $A$ alleles at $y$ multiplied by the probability of dispersing from $y$ to $x$;
looking in the other direction, this is also the probability that an $A$ allele at $x$ had a parent at $y$.
In other words, lineages move as the random walk determined by dispersal,
but biased towards areas that produce more of the selected allele that they carry.
Making the same assumptions underlying equation \eqref{eqn:cline_pde},
we show in Appendix \ref{apx:lineage_derivation}
that the lineage of an $A$ allele moves as Brownian motion with speed $\sigma$
with drift $s \grad \log p(x,t)$ (i.e., it is pushed uphill on $\log p$ at speed $s$).
Since $p=1$ far from the zone to the left, 
and is proportional to $\exp(-x\sqrt{s}/\sigma)$ to the right,
then roughly, lineages on ``their own'' side wander randomly,
while lineages on ``the wrong'' side are pushed at constant speed $\sigma/\sqrt{s}$ 
back towards the side where they are more common.
Since an $A$ allele must by definition have been inherited from the $A$ side of the barrier 
at the time of secondary contact,
this push must get more intense the closer it is to the time of secondary contact.
Lineages of $B$ alleles do the same thing, of course, but in the opposite direction.

\paragraph{Lineages at linked loci}
The behavior of a lineage at a linked locus is similar,
except that when in heterozygotes the lineage may move between backgrounds:
if we follow back through time the lineage of a locus we find linked to an $A$ allele, 
it will first tend to be inherited from ancestors to the left (as $A$ lineages drift to the left),
but if it spends time in the clinal region where recombination may happen,
it may be found to have been inherited from a $B$-carrying individual,
whose ancestors will tend to be more from the right.
Suppose we sample an allele today at a locus at recombination distance $r$ to the selected site.
If $X_t$ is the location of its ancestor $t$ generations ago,
and $Z_t$ is the identity of the selected allele that ancestor carried,
then we say that $X$ moves as a diffusion pushed by either $\log(p)$ or $\log(1-p)$,
and that $Z$ jumps between $A$ and $B$ at rate either $r p$ or $r(1-p)$,
depending on the identity of $Z$.
In symbols,
\begin{align}
    \begin{aligned} \label{eqn:lineage_motion}
        dX_t &= \begin{cases}
             s \grad \log(p(X_t,t)) dt + dB_t \qquad & \text{if } Z_t = A \\
             s \grad \log(1-p(X_t,t)) dt + dB_t \qquad & \text{if } Z_t = B 
        \end{cases} \\
        \P\{ Z_{t+\epsilon} = B \given Z_t = A \} &= \epsilon r (1-p(X_t,t)) + O(\epsilon^2) \\
        \P\{ Z_{t+\epsilon} = A \given Z_t = B \} &= \epsilon r p(X_t,t) + O(\epsilon^2)  .
    \end{aligned}
\end{align}

\textbf{figure of a lineage moving back and forth?}

\paragraph{How fast do linked clines relax?}
Although loci not under selection can in principle spread across the cline unimpeded,
in practice it can take quite some time for even unlinked neutral loci to homogenize,
due to the decreased fitness of heterozygotes \citep{bartonbarrier}
and the relative slowness of diffusive movement \citep{Sedghifar2015}.
Suppose we sample across at cline $\tau$ units of time after secondary contact.
How far away from the selected locus do we expect to see clines that have begun to flatten out more
than the selected cline?
A linked cline will have begun to flatten out if there is an appreciable chance that
a lineage at the linked locus initially linked to an $A$ allele 
at some point has recombined from an individual carrying the $B$ allele.
This implies that loci closer than $1/\tau$ to the selected site
certainly won't have a substantially wider cline than the selected cline.
\plr{Figure showing this? 
    For instance, either: (a) time-genome heatmap of number of heterozygotes (from PDE);
    or (b) a few time-space heatmaps of frequencies at different values of r (from PDE);
    or (c) a few space-genome heatmaps of frequencies at different times (from sims)
}
Of course, this calculation assumes that every generation provides an opportunity for recombination 
onto the other background,
but only recombinations in heterozygotes will move the linked lineage between selected backgrounds.
Since heterozygotes for the selected allele are only found at high frequency in the cline,
and lineages are generally pushed away from the cline but have no bias far away,
the amount of time a lineage spends in heterozygotes should grow as $\sqrt{\tau}$ for large $\tau$,
and so the width of the genomic region showing clines about the selected locus could be 
substantially larger than $1/\tau$.
However, this distinction appears hard to observe for realistic parameter values.

\paragraph{PDE for linked clines}
Clines at linked sites are given by
the probability that 
an allele sampled $t$ generations after secondary contact at location $x$,
at recombination distance $r$ to a selected allele of type $z$
is inherited from an individual of ancestry $A$,
where $z$ can be either $A$ or $B$.
Denote this probability $q_z(x,t,r)$.
In other words,
\begin{align}
    q_z(x,t,r) = \P^x \{Z_t = A\} .
\end{align}
The description above implies that $q_r$ solves the equation
\begin{align}
    \begin{aligned}  \label{eqn:q_pde}
    \dot q_A(x,t,r) 
            &= \grad \log(p(x,t)) \cdot \grad q_A(x,t,r) 
                + \frac{\sigma^2}{2} \Delta q_A(x,t,r) 
            \\ &\qquad {} + 
                r (1-p(x,t))(q_B(x,t,r)-q_A(x,t,r))  \\
    \dot q_B(x,t,r) &= \grad \log(1-p(x,t)) \cdot \grad q_B(x,t,r) 
            + \frac{\sigma^2}{2} \Delta q_B(x,t,r)
            \\ &\qquad {} + 
            + r (1-p(x,t))(q_A(x,t,r)-q_B(x,t,r))  ,
    \end{aligned} \\
\end{align}
with boundary conditions $q_A(x,0,r)=1$ and $q_B(x,0,r)=0$.
We will have more use for the differential operators on the right-hand sides of these equations,
so write these also as 
$\dot q_A = G_r^A q_A + r (1-p) (q_A-q_B)$
and 
$\dot q_B = G_r^B q_B + r (1-p) (q_B-q_A)$.

\paragraph{Numerical computation}
To determine how the cline at a linked locus is expected to relax,
we can solve the partial differential equations \eqref{eqn:cline_pde} and \eqref{eqn:q_pde} numerically,
producing, for instance, a heatmap of expected local frequencies of $A$ ancestry
across space and time, as in Figure \ref{fig:pde_clines}.
Shown is the expected frequency of ancestry $A$ at location $x$ and time $t$ at a site at recombination distance $r$ from the selected site,
which we denote $p(x,t,r)$ and can compute as $p(x,t,r) = p(x,t) q_A(x,t,r) + (1-p(x,t)) q_B(x,t,r)$.

@@

\paragraph{Haplotype lengths}
We can also find the probability of seeing an entire block of genome (haplotype)
of a single ancestry at a given location and time.
As in \citet{Sedghifar2015},
to do this we need only keep track of both lineages when a recombination in the middle of a block occurs,
so that lineages will act like a particular class of labeled, branching diffusions,
where the total branching rate is conserved.
We will not consider subsequent coalescence.
The general description of the process is as follows.
The lineage of a segment of genome behaves as a linked locus described in equations \eqref{eqn:lineage_motion},
with recombination distance $r$ equal to the rate at which the segment recombines away from the selected site.
(If the selected site lies inside the segment, $r=0$.)
Additionally, at rate equal to the genetic length of the segment,
a recombination occurs, at which point the segment splits into two segments at a uniformly chosen location,
each of which procedes as before, independently.

\paragraph{Spatial haplotype patterns}
Furthermore, we expect that stretches of $A$ ancestry will tend to be shorter the further one moves into the $B$ side of the cline,
because they have had more opportunities to recombine with $B$ haplotypes.
However, stretches of $A$ ancestry that contain the selected site are expected to be longer
than those that do not at the same spatial location,
because lineages containing the selected site must have been inherited from the $A$ side of the cline recently --
concretely, they move at speed roughly $\sigma/\sqrt{s}$,
so we expect (selected) $A$ alleles at distance $x$ from the cline center
to have last had an ancestor on the $A$ side of the cline around $x \sqrt{s}/\sigma$ generations ago,
and so the enclosing $A$ haplotype is no longer than $\sigma/(x \sqrt{s})$.
\plr{(check predicted decay of mean length $1/x$ in simulations)}
For linked haplotypes not containing the selected site, the situation is more complicated, and harder to intuit.

\paragraph{General haplotype probabilities}
To quantify haplotype abundances,
we'd like to know the probability that the entire segment of genome between positions $a$ and $b$ relative to the selected site
in an individual sampled at spatial location $x$ and time $t$ after the cline has formed
is of ancestry $A$.
Writing this probability $g_A(x,t;a,b)$,
branching adds a term to equation \eqref{eqn:q_pde}, so that 
\plr{write this equation more compactly?}
\begin{align}
    \begin{aligned} \label{eqn:g_pde}
        \dot g_A(x,t;a,b) 
            &= G_{r(a,b)}^A g_A(x,t;a,b) 
            - (b-a) g_A(x,t;a,b) 
            \\ {} & \qquad 
            + p(x,t) \int_a^b {
                g_A(x,t;a,\theta) g_A(x,t;\theta,b) 
                - g_A(x,t;a,b) .
            } d\theta
            \\ {} & \qquad 
            + (1-p(x,t)) \int_a^0 {
                g_B(x,t;a,\theta) g_A(x,t;\theta,b)
            } d\theta
            \\ {} & \qquad 
            + (1-p(x,t)) \int_0^b {
                g_A(x,t;a,\theta) g_B(x,t;\theta,b)
            } d\theta,
    \end{aligned}
\end{align}
where the integral $\int_a^0$ is zero if $a>0$ and the integral $\int_0^b$ is zero if $b<0$;
and $r(a,b)$ is the distance from the segment to the selected site:
$r(a,b)=0$ if $ab<0$ and $r(a,b)=\min(|a|,|b|)$ otherwise.
The equation for $g_B(x,t;a,b)$ is identical after exchanging $A \leftrightarrow B$ and
$p \leftrightarrow (1-p)$.
The boundary conditions are $g_A(x,0;a,b)=1$ and $g_B(x,0;a,b)=0$.

@@

We implemented forwards-time simulations of diploid individuals on a one-dimensional row of demes
with nonoverlapping generations and fixed population sizes (a Wright-Fisher model),
and a single locus with selection against heterozygotes in the center of one of two chromosomes, each of length 1 Morgan.
Each deme has exactly $n_d$ diploid individuals at the start of each generation.
Each generation,
first, each individual migrated to a new deme offset by $\lfloor Z + 1/2 \rfloor$ from their original position,
where $Z$ is a Gaussian random variable with mean zero and variance $\sigma^2$.
(The mean displacement is zero, and when comparing to theory we compute $\sigma$ as the standard deviation of this distribution.)
Migrants past either end of the range remained at the ends.
Then, 
fitnesses were computed (heterozygotes have fitness $1-s$; homozygotes have fitness 1),
and in each deme $n_d$ pairs of parents are chosen, with replacement,
with probability proportional to their fitness.
Selfing is allowed.
Since migration is not conservative, demes may have no available parents;
in this case, parents are chosen from other demes with probability proportional to fitness
multiplied by $\exp(-x^2/\sigma^2)$, where $x$ is the distance to the other deme.
The next generation is formed by carrying out meiosis in each parent
and combining the resulting gametes.
Meiosis results in alternating blocks of the gamete's chromsome
being inherited from the two parental chromosomes,
with the blocks separated by a Poisson(1) number of uniformly chosen recombination points along the chromosome,
and the order of the parental chromosomes chosen randomly.
The simulation software works by recording for each chromosome
a list of ancestry breakpoints and the index of the ancestor in the first generation
that the chromosome inherited that region of genome from;
genotyping at individual loci was performed by looking up the ancestor's genotype.

Hybrid zones of ages 10, 100 and 1000 generations were simulated under conditions of $\sigma=1$ and $s=0.1$. From this we measure unbroken tract lengths to the right of the selected locus. These were compared to the output of neutral simulations with $s=0$

>>>>

\section{Introduction}
The process of speciation commonly involves allopatric divergence in relative isolation \citep{Coyne2004}. When formerly isolated populations come back into contact \yb{the exchange of genes across a geographic hybrid zone can establish a gradient of ancestry across geography, \citep{Barton1985}}. In addition to steep clines in ancestry, recently formed hybrid zones are characterized by a large physical scale of linkage disequilibrium in ancestry (ancestry-LD). If genetic variation can be exchanged between species via hybridization, both of these signals diminish over time throughout the genome; however, subsets of the genome that are under selection may be expected to maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}. The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986}. \yb{It is clear that selection against admixed ancestry at one locus will influence patterns of ancestry at physically linked loci, however the expected length of ancestral tracts around selected loci in admixing populations is poorly understood.}

Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place every generation. The expected distribution of lengths of unbroken ancestry have been previously described under neutral conditions \cite[e.g.][]{Gravel2012,Sedghifar2015}.  
Here, we investigate expected patterns of co-ancestry \yb{ --- that is, the genomic scale at which ancestry strongly covaries between sites --- } across a hybrid zone in the presence of selection. 
Selection against migrant \yb{alleles} in a hybrid zone results in reduced frequencies of foreign ancestry surrounding targets of selection. 
The effect on patterns of ancestral block lengths is, however, less well understood. 
Because of rapid removal from the population, we expect to observe \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone. \alisa{While empirical patterns of block length in hybrid zones have not been well studied, the availability of dense genotype data in combination with linkage information pave a way to utilize such patterns to better understand the processes operating in hybrid zones. }

While population genetic models of hybrid zones have been extensively explored, these have typically focused on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}. 
These models predict cline width to be narrowest at loci close to selected loci, and cline width at less tightly linked loci to be wider and more heavily influenced by hybrid zone age and rates of gene flow. 
\yb{The focus on cline width provide an expectation for how selection shapes the covariation in ancestry at a single locus over \emph{geographic space}, our aim is to provide an expectation for how selection shapes the covariation in ancestry at loci over \emph{genomic space.}}  
%As denser genotype data becomes more readily available, genome-wide patterns of linkage can be utilized in the analysis of gene hybrid zones.

@@

\paragraph{Lineages at the selected locus}
Now suppose that the frequency profile of the selected allele $p(x,t)$ is given.
We will understand how lineages of sampled individuals behave,
both at and away from the selected site,
by following lineages as they move between backgrounds,
as done for instance in \citet{selectioncoal,ralph2015patchy}.
Consider the collection of $A$ alleles found at location $x$.
The expected number of such alleles produced by location $y$ is proportional to 
the number of $A$ alleles at $y$ multiplied by the probability of dispersing from $y$ to $x$;
looking in the other direction, this is also the probability that an $A$ allele at $x$ had a parent at $y$.
In other words, lineages move as the random walk determined by dispersal,
but biased towards areas that produce more of the selected allele that they carry.
Making the same assumptions underlying equation \eqref{eqn:cline_pde},
we show in Appendix \ref{apx:lineage_derivation}
that the lineage of an $A$ allele moves as Brownian motion with speed $\sigma$
with drift $s \grad \log p(x,t)$ (i.e., it is pushed uphill on $\log p$ at speed $s$) \alisa{note to self to read through appendix when written}.
Since $p=1$ far from the zone to the left, 
and is proportional to $\exp(-x\sqrt{s}/\sigma)$ to the right,
then roughly, lineages on ``their own'' side wander randomly,
while lineages on ``the wrong'' side are pushed at constant speed $\sigma/\sqrt{s}$ 
back towards the side where they are more common.
Since an $A$ allele must by definition have been inherited from the $A$ side of the barrier 
at the time of secondary contact,
this push must get more intense the closer it is to the time of secondary contact.
Lineages of $B$ alleles do the same thing, of course, but in the opposite direction.

\paragraph{Lineages at linked loci}
The behavior of a lineage at a linked locus is similar,
except that when in heterozygotes the lineage may move between backgrounds:
if we follow back through time the lineage of a locus we find linked to an $A$ allele, 
it will first tend to be inherited from ancestors to the left (as $A$ lineages drift to the left),
but if it spends time in the clinal region where recombination may happen,
it may be found to have been inherited from a $B$-carrying individual,
whose ancestors will tend to be more from the right.
Suppose we sample an allele today at a locus at recombination distance $r$ to the selected site.
If $X_t$ is the location of its ancestor $t$ generations ago,
and $Z_t$ is the identity of the selected allele that ancestor carried,
then we say that $X$ moves as a diffusion pushed by either $\log(p)$ or $\log(1-p)$,
and that $Z$ jumps between $A$ and $B$ at rate either $r p$ or $r(1-p)$,
depending on the identity of $Z$ \alisa{is $t$ the time of recombination?}.
In symbols,
\begin{align}
    \begin{aligned} \label{eqn:lineage_motion}
        dX_t &= \begin{cases}
             s \grad \log(p(X_t,t)) dt + dB_t \qquad & \text{if } Z_t = A \\
             s \grad \log(1-p(X_t,t)) dt + dB_t \qquad & \text{if } Z_t = B 
        \end{cases} \\
        \P\{ Z_{t+\epsilon} = B \given Z_t = A \} &= \epsilon r (1-p(X_t,t)) + O(\epsilon^2) \\
        \P\{ Z_{t+\epsilon} = A \given Z_t = B \} &= \epsilon r p(X_t,t) + O(\epsilon^2)  .
    \end{aligned}
\end{align}

\textbf{figure of a lineage moving back and forth?} \alisa{A conceptual cartoon of this process is shown in Figure~\ref{Fig:schematic}.}

\paragraph{How fast do linked clines relax?}
Although loci not under selection can in principle spread across the cline unimpeded,
in practice it can take quite some time for even unlinked neutral loci to homogenize,
due to the decreased fitness of heterozygotes \citep{bartonbarrier}
and the relative slowness of diffusive movement \citep{Sedghifar2015}.
Suppose we sample across at cline $\tau$ units of time after secondary contact.
How far away from the selected locus do we expect to see clines that have begun to flatten out more
than the selected cline?
A linked cline will have begun to flatten out if there is an appreciable chance that
a lineage at the linked locus initially linked to an $A$ allele 
at some point has recombined from an individual carrying the $B$ allele.
This implies that loci closer than $1/\tau$ to the selected site
certainly won't have a substantially wider cline than the selected cline.
\plr{Figure showing this? 
    For instance, either: (a) time-genome heatmap of number of heterozygotes (from PDE);
    or (b) a few time-space heatmaps of frequencies at different values of r (from PDE);
    or (c) a few space-genome heatmaps of frequencies at different times (from sims)
}
Of course, this calculation assumes that every generation provides an opportunity for recombination 
onto the other background,
but only recombinations in heterozygotes will move the linked lineage between selected backgrounds.
Since heterozygotes for the selected allele are only found at high frequency in the cline,
and lineages are generally pushed away from the cline but have no bias far away,
the amount of time a lineage spends in heterozygotes should grow as $\sqrt{\tau}$ for large $\tau$,
and so the width of the genomic region showing clines about the selected locus could be 
substantially larger than $1/\tau$.
However, this distinction appears hard to observe for realistic parameter values. \alisa{perhaps move some of this to discussion?}

\paragraph{PDE for linked clines}
Clines at linked sites are given by
the probability that 
an allele sampled $t$ generations after secondary contact at location $x$,
at recombination distance $r$ to a selected allele of type $z$
is inherited from an individual of ancestry $A$,
where $z$ can be either $A$ or $B$.
Denote this probability $q_z(x,t,r)$.
In other words,
\begin{align}
    q_z(x,t,r) = \P^x \{Z_t = A\} .
\end{align}
The description above implies that $q_r$ \alisa{not $q_z$?} solves the equation
\begin{align}
    \begin{aligned}  \label{eqn:q_pde}
    \dot q_A(x,t,r) 
            &= \grad \log(p(x,t)) \cdot \grad q_A(x,t,r) 
                + \frac{\sigma^2}{2} \Delta q_A(x,t,r) 
            \\ &\qquad {} + 
                r (1-p(x,t))(q_B(x,t,r)-q_A(x,t,r))  \\
    \dot q_B(x,t,r) &= \grad \log(1-p(x,t)) \cdot \grad q_B(x,t,r) 
            + \frac{\sigma^2}{2} \Delta q_B(x,t,r)
            \\ &\qquad {} + 
            + r (1-p(x,t))(q_A(x,t,r)-q_B(x,t,r))  ,
    \end{aligned} \\
\end{align}
with boundary conditions $q_A(x,0,r)=1$ and $q_B(x,0,r)=0$.
We will have more use for the differential operators on the right-hand sides of these equations,
so write these also as 
$\dot q_A = G_r^A q_A + r (1-p) (q_A-q_B)$
and 
$\dot q_B = G_r^B q_B + r (1-p) (q_B-q_A)$.
\alisa{A few words about what $G$ is?}

\paragraph{Numerical computation}
To determine how the cline at a linked locus is expected to relax,
we can solve the partial differential equations \eqref{eqn:cline_pde} and \eqref{eqn:q_pde} numerically,
producing, for instance, a heatmap of expected local frequencies of $A$ ancestry
across space and time, as in Figure \ref{fig:pde_clines} \alisa{which shows}
 the expected frequency of ancestry $A$ at location $x$ and time $t$ at a site at recombination distance $r$ from the selected site,
\alisa{denoted} $p(x,t,r)$, and can be computed as $p(x,t,r) = p(x,t) q_A(x,t,r) + (1-p(x,t)) q_B(x,t,r)$. \alisa{Some details on numerical solution (e.g. R package?)}

@@

\paragraph{Haplotype lengths}
We can also find the probability of seeing an entire block of genome (haplotype)
of a single ancestry at a given location and time.
As in \citet{Sedghifar2015},
to do this we need only keep track of both lineages when a recombination in the middle of a block occurs (Figure~\ref{Fig:schematic}),
so that lineages will act like a particular class of labeled, branching diffusions,
where the total branching rate is conserved.
We will not consider subsequent coalescence.
The general description of the process is as follows.
The lineage of a segment of genome behaves as a linked locus described in equations \eqref{eqn:lineage_motion},
with recombination distance $r$ equal to the rate at which the segment recombines away from the selected site.
(If the selected site lies inside the segment, $r=0$.)
Additionally, at rate equal to the genetic length of the segment,
a recombination occurs, at which point the segment splits into two segments at a uniformly chosen location,
each of which proceeds as before, independently.

\paragraph{Spatial haplotype patterns}
We expect that\alisa{, on average,} stretches of $A$ ancestry will tend to be shorter the further one moves into the $B$ side of the cline,
because they have had more opportunities to recombine with $B$ haplotypes.
However, stretches of $A$ ancestry that contain the selected site are expected to be longer
than those that do not at the same spatial location,
because lineages containing the selected site must have been inherited from the $A$ side of the cline recently --
concretely, they move at speed roughly $\sigma/\sqrt{s}$,
so we expect (selected) $A$ alleles at distance $x$ from the cline center
to have last had an ancestor on the $A$ side of the cline around $x \sqrt{s}/\sigma$ generations ago,
and so the enclosing $A$ haplotype is no longer than $\sigma/(x \sqrt{s})$.
\plr{(check predicted decay of mean length $1/x$ in simulations)}
For linked haplotypes not containing the selected site, the situation is more complicated, and harder to intuit.

\paragraph{General haplotype probabilities}
To quantify haplotype abundances,
we'd like to know the probability that the entire segment of genome between positions $a$ and $b$ relative to the selected site
in an individual sampled at spatial location $x$ and time $t$ after the cline has formed
is of ancestry $A$.
Writing this probability $g_A(x,t;a,b)$,
branching adds a term to equation \eqref{eqn:q_pde}, so that 
\plr{write this equation more compactly?}
\begin{align}
    \begin{aligned} \label{eqn:g_pde}
        \dot g_A(x,t;a,b) 
            &= G_{r(a,b)}^A g_A(x,t;a,b) 
            - (b-a) g_A(x,t;a,b) 
            \\ {} & \qquad 
            + p(x,t) \int_a^b {
                g_A(x,t;a,\theta) g_A(x,t;\theta,b) 
                - g_A(x,t;a,b) .
            } d\theta
            \\ {} & \qquad 
            + (1-p(x,t)) \int_a^0 {
                g_B(x,t;a,\theta) g_A(x,t;\theta,b)
            } d\theta
            \\ {} & \qquad 
            + (1-p(x,t)) \int_0^b {
                g_A(x,t;a,\theta) g_B(x,t;\theta,b)
            } d\theta,
    \end{aligned}
\end{align}
where the integral $\int_a^0$ is zero if $a>0$ and the integral $\int_0^b$ is zero if $b<0$;
and $r(a,b)$ is the distance from the segment to the selected site:
$r(a,b)=0$ if $ab<0$ and $r(a,b)=\min(|a|,|b|)$ otherwise.
The equation for $g_B(x,t;a,b)$ is identical after exchanging $A \leftrightarrow B$ and
$p \leftrightarrow (1-p)$.
The boundary conditions are $g_A(x,0;a,b)=1$ and $g_B(x,0;a,b)=0$.
\alisa{Is the selected site at $\theta=0$?}

@@

We implemented forwards-time simulations of diploid individuals on a one-dimensional row of demes
with non-overlapping generations and fixed population sizes (a Wright-Fisher model). Individuals are diploid, with haploid number $n=2$ chromosomes, each of length 1 Morgan. One chromosome pair harbors in its center (at position 0.5M) a single locus that reduces fitness by $s$ in heterozygotes, while the other contains no sites under direct selection.

Each deme has exactly $N_d$ diploid individuals at the start of each generation.
At the start of each generation, every individual migrates to a new deme  offset by $\lfloor Z + 1/2 \rfloor$ from their original position,
where $Z$ is a Gaussian random variable with mean zero and variance $\sigma^2$.
(The mean displacement is zero, and when comparing to theory we compute $\sigma$ as the standard deviation of this distribution.)
Migrants past either end of the range remained at the terminal demes.
Then, 
fitnesses were computed (heterozygotes have fitness $1-s$; homozygotes have fitness 1),
and in each deme $N_d$ pairs of parents are chosen, with replacement,
with probability proportional to their fitness and selfing allowed.
Since migration is not conservative, demes may have no available parents;
in this case, parents are chosen from other demes with probability proportional to fitness
multiplied by $\exp(-x^2/\sigma^2)$, where $x$ is the distance to the other deme.
The next generation is formed by carrying out meiosis in each parent
and combining the resulting gametes such that each pair of parents leaves one descendant.
Meiosis results in alternating blocks of the gamete's chromsome
being inherited from the two parental chromosomes,
with the blocks separated by a Poisson(1) number of uniformly chosen recombination points along the chromosome,
and the order of the parental chromosomes chosen randomly.
The simulation software works by recording for each chromosome
a list of ancestry breakpoints and the index of the ancestor in the first generation
that the chromosome inherited that region of genome from;
genotyping at individual loci was performed by looking up the ancestor's genotype.

Hybrid zones of ages 10, 100 and 1000 generations \alisa{these numbers may need to be revised?} were simulated under conditions of $\sigma=1$ and $s=0.1,0.01$. From this we measure the length of unbroken tracts of ancestry $B$ surrounding a given position along a chromosome (if the locus in question was of ancestry $A$ the ancestry block had length $0$). These were compared to the output of neutral simulations with $s=0$
